Appendix: Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model

### 2020-08-25 Tex/PA_appendix_rr/appendix.tex
included software information in main and appendix and explained SA algorithm in more detail

The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant. However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore. Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$. But even this approach started to fail for networks around the turn of the millennium. Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}. 
The MCMLE for these large citation networks was obtained by setting the starting value according to an approach introduced by \citet{SchmidHunter2020}. This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$. However, the MPLE of networks with the same sufficient statistics is not necessarily the same. Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a network $C_t^*$ using simulated annealing \citep{Kirkpatrick83}, with $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$ and a weak dependence structure among unfixed ties. The MPLE of such a network is similar to the MLE, making it an effective starting value for the MCMLE algorithm. Since for some networks the MCMLE was only obtainable using simulated annealing starting value, the final results in the paper have all been estimated using simulated annealing.

>>>>

The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant. However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore. Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$. But even this approach started to fail for networks around the turn of the millennium. Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}. 
The MCMLE for these large citation networks was obtained by setting the starting value according to an novel approach introduced by \citet{SchmidHunter2020}. This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$. However, the MPLE of networks with the same sufficient statistics is not necessarily the same. Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a new network $C_t^*$ on the same nodes as the observed network that satisfies $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$, and has a weak dependence structure among unfixed ties. Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}. For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE. In addition, the same sufficient statistic between $C_t^*$ and the observed network guarantees the same MLE between the two networks. This makes the new network's MPLE an effective starting value for the MCMLE algorithm. Since for some networks the MCMLE was only obtainable using simulated annealing starting value, the final results in the paper have all been estimated using simulated annealing.
The simulated annealing algorithm for finding an improved starting value for c-ERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.
